Multi-fluids for Representing Sub-filter-scale Convection

### 2020-04-21 submit2/manuscript.tex
Will's changes and Hilary's changes to figures

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%
\addeditor{HW}

@@

\citeA{TWV+18} proposed multi-fluid modelling of convection in order
to treat non-equilibrium and net mass transport \remove[HW]{correctly and} consistently.
\add[HW]{The multi-fluids approach has an advantage over ADHOC and CLUBB in solving prognostic equations for first-order moments directly (means of eg. updraft and environment) rather than solving prognostic equations for higher-order moments of the whole fluid in order to diagnose means of updrafts and the environment. Solving directly for sub-fluid means should be less sensitive to closure assumptions.}
\remove{with the same numerics and physics to prognose properties of separate parts of the fluid, for example updrafts, the stable environment and downdrafts dynamics}. \citeA{Yano14} derived similar
equations \remove[HW]{for multiple segments} and showed how this is a generalisation
of the mass-flux formulation. The \citeA{TKP+18} ``extended eddy
diffusivity mass flux'' or extended EDMF scheme \remove[HW]{also} uses similar
equations but \citeA{TKP+18}
\change[HW]{didn't propose a suitable numerical method for solving the equations.}
{only solve the equations in a single column without coupling to the continuity equation of a dynamical core.}
\change[HW]{Multi-fluids promises big improvements
in the representation of sub-grid-scale convection and unification
with the representation of turbulence, particularly at the grey zone.
However, there is a long way to go, particularly as multi-fluids}
{The disadvantage of the multi-fluids approach is that it} requires
a dedicated model rather than being a stand alone parameterisation
for a dynamical core. \citeA{TEB19} and \citeA{WM19} presented solutions
of \change[HW]{the multi}{two}-fluid equations
\change[HW]{assuming two fluids that share}{using} the same pressure
\add[HW]{for both fluids}.
\remove[HW]{(or that pressure differences are predominantly due to drag
between the fluids)}. \citeA{TEB19} showed that these equations \remove[HW]{(without
drag)} are in fact unstable and can be stabilised by diffusion of vertical
velocity whereas \citeA{WM19} stabilised the equations with diffusion
and drag between fluids. Neither of these stabilisation techniques
are acceptable;
\add[HW]{sufficient diffusion to stabilise the multi-fluid equations may smooth out real features of the flow and the diffusion between fluids and drag used by \protect\citeA{WM19} meant that both fluids tended to move together rather than through each other.}
In this paper we show that there are significant
and sustained pressure differences between the fluids (section \ref{sec:results}) \add[HW]{and that the use of separate pressures stabilises the equations}.

@@

Many of the challenges of representing convection with the multi-fluid
Navier-Stokes equations \cite<e.g. those described in>{WM19} carry
over to the Boussinesq equations \add[HW]{because} convection is buoyancy dominated
and close to incompressible. The resolved test cases presented in
section \ref{sec:results} give very similar results for Navier-Stokes
and Buossinesq equations. Therefore for simplicity we focus on the
\remove[HW]{multi-fluid} Boussinesq equations without background stratification:
\note[HW]{This is additional material until the start of 2.1}
\begin{eqnarray}
\frac{\partial \mathbf{u}}{\partial t} + \nabla\cdot(\mathbf{u}\mathbf{u})
+\nabla P & = & b \mathbf{k}+\nu\nabla^{2}\mathbf{u}
\label{eq:singleMom}
\\
\frac{\partial b}{\partial t} + \nabla\cdot(\mathbf{u}b)
 & = & \alpha\nabla^{2}b
\label{eq:singleb}
\\
\nabla\cdot\mathbf{u} & = & 0
\label{eq:singleDivFree}
\end{eqnarray}
where variables are defined in table \ref{tab:defns}.

\begin{table}
\begin{tabular}{c>{\raggedright}p{0.85\textwidth}}
$\sigma_{i}$ & Volume fraction of fluid $i$ so that $\sum_{i}\sigma_{i}=1$
\tabularnewline
\add[HW]{$\mathbf{u}$} & 
\add[HW]{Velocity of the single-fluid equations  ($\text{m}\text{s}^{-1}$) and velocity averaged over all fluids of the multi-fluid equations}
\tabularnewline
$\mathbf{u}_{i}$ & Velocity of fluid $i$ ($\text{m}\text{s}^{-1}$)
\tabularnewline
$w_{i}$ & Vertical component of $\mathbf{u}_{i}$ ($\text{m}\text{s}^{-1}$)\tabularnewline
$S_{ij}$ & Transfer from fluid $i$ to $j$ ($\text{s}^{-1}$)
\tabularnewline
\add[HW]{$P$} & \add[HW]{Non-hydrostatic pressure of the single-fluid equations $=p^{\prime}/\rho_{r}$ ($\text{m}^{2}\text{s}^{-2}$) and pressure averaged over all fluids of the multi-fluid equations}
\tabularnewline
$P_i$ & Non-hydrostatic pressure of fluid $i$
\tabularnewline
\add[HW]{$p_i$} & \add[HW]{$P_i - P$}
\tabularnewline
$\psi_{ij}^{T}$ & Value of variable $\psi$ transferred from fluid $i$ to $j$
\tabularnewline
$b$ & Buoyancy of single-fluid equations defined as -$g\rho^{\prime}/\text{\ensuremath{\rho}}_{r}$
where $\rho^{\prime}$ are departures in density from a horizontally
uniform reference, $\rho_{r}$ ($\text{m}\text{s}^{-2}$) \add[HW]{and buoyancy averaged over all fluids of the multi-fluid equations}
\tabularnewline
$b_{i}$ & Buoyancy of fluid $i$ ($\text{m}\text{s}^{-2}$)
\tabularnewline
$D_{i}\big/Dt$ & Total derivative with respect to fluid $i$ $=\partial/\partial t+\mathbf{u}_{i}\cdot\nabla$
\tabularnewline
$\mathbf{D}_{ij}$ & Drag on fluid $i$ from fluid $j$ ($\text{m}\text{s}^{-2}$)
\tabularnewline
$\gamma$ & \change[HW]{Coefficient for setting the pressure local for each fluid}{Compressibility for calculating $p_i$} ($\text{m}^{2}\text{s}^{-1}$)
\tabularnewline
$C_{D}$ & Drag coefficient $=0$ or 1
\tabularnewline
$r_{c}$ & Plume radius used for defining the drag between fluids (m)\tabularnewline
$\alpha$ & Diffusivity of buoyancy ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
$\nu$ & Viscosity ($\text{m}^{2}\text{s}^{-1}$)
\tabularnewline
\end{tabular}
\caption{Definitions of variables for the single- and multi-fluid Boussinesq equations.\label{tab:defns}}
\end{table}

The single-fluid Boussinesq equations will be conditionally averaged following \citeA{TWV+18} to give the multi-fluid Boussinesq equations. Lagrangian labels for each fluid $i$ are defined as $I_i=1$ in fluid $i$ and zero elsewhere. 
For the derivation of the conditionally averaged equations we will not consider relabelling of fluid \add[HW]{and so} $I_i$ satisfies:
\begin{equation}
\frac{DI_i}{Dt}=0.
\label{eq:LagrangianLabel}
\end{equation}
Applying conditional volume average to $I_i$ we define the volume fraction of fluid $i$ and the velocity and buoyancy of fluid $i$:
\begin{eqnarray}
\sigma_i &=& \overline{I_i} \\
\sigma_i \mathbf{u}_i &=& \overline{I_i \mathbf{u}} \\
\sigma_i b_i &=& \overline{I_i b}
\label{eq:defineFluidFields}
\end{eqnarray}
which implies
\begin{equation}
\sum_{i}\sigma_{i}  =  1.
\label{eq:sumOne}
\end{equation}
As we are considering the Boussinesq equations we assume that variations in density only affect the buoyancy terms so $\mathbf{u}_i$ and $b_i$ are defined without density weighting.

To derive a transport equation for $\sigma_i$ we combine (\ref{eq:LagrangianLabel}) with $I_i$ times (\ref{eq:singleDivFree}) and volume average. After permuting operators with the averaging \cite<as described by>{TWV+18} and using the definitions of $\sigma_i$ and $\mathbf{u}_i$ we get the evolution equation for $\sigma_i$:
\begin{equation}
\frac{\partial\sigma_{i}}{\partial t}+\nabla\cdot\sigma_{i}\mathbf{u}_{i}  =  \sum_{j\ne i}\left\{ \sigma_{j}S_{ji}-\sigma_{i}S_{ij}\right\}
\label{eq:sigma}
\end{equation}
which also includes transfers between fluids with $\sigma_{i}S_{ij}$ being the transfer from fluid $i$ to $j$.

Next we will consider conditionally averaging the terms of the momentum (\ref{eq:singleMom}) and buoyancy (\ref{eq:singleb}) equations. Conditionally averaging the advection terms leads to sub-filter-scale fluxes, $F_{SF}$, since the velocity and buoyancy are not uniform within each fluid at the filter scale:
\begin{eqnarray}
\overline{I_i \nabla \cdot (\mathbf{u} \mathbf{u})} &=& \nabla \cdot (\sigma_i \mathbf{u}_i  \mathbf{u}_i) + F_\text{SF}^{\mathbf{u}_i} \\
\overline{I_i \nabla (\mathbf{u} b)} &=&  \nabla \cdot (\sigma_i \mathbf{u}_i b_i) + F_\text{SF}^{b_i}
\label{eq:filterAdvection}
\end{eqnarray}
The sub-filter-scale fluxes will be absorbed into the diffusion terms for momentum and buoyancy.
%
\note[HW]{Dan, can we do better than this?}
Lots of terms appear when attempting to conditionally average the diffusion terms which we do not know how to parameterise. Therefore we will use simply:
\begin{eqnarray}
\overline{I_i \nu \nabla^2 \mathbf{u}} &=& \sigma_i \nu \nabla^2 \mathbf{u}_i \\
\overline{I_i \alpha \nabla^2 b} &=& \sigma_i \alpha \nabla^2b_i .
\end{eqnarray} 
%
Unlike \citeA{TWV+18} we define separate pressures for each fluid, $P_i$. The pressures are defined to satisfy:
\begin{equation}
\overline{I_i \nabla P} = \sigma_i \nabla P_i.
\end{equation}
%
Combining these gives the conditionally averaged momentum and buoyancy equations:
\begin{eqnarray}
\frac{\partial\sigma_i\mathbf{u}_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} \mathbf{u}_{i})
+ \sigma_i \nabla P_{i}
& = &
\sigma_i b_{i}\mathbf{k}
+
\nu\sigma_i\nabla^{2}\mathbf{u}_{i}
+
\sum_{j\ne i}\left\{ \sigma_{j}S_{ji}\mathbf{u}_{ji}^{T}
 - \sigma_{i}\mathbf{D}_{ij}\right\}
\label{eq:momFlux}\\
\frac{\partial\sigma_i b_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} b_{i})
& = &
\alpha\sigma_i\nabla^{2}b_{i}
+
\sum_{j\ne i} \sigma_{j}S_{ji} b_{ji}^{T}.
\label{eq:bFlux}
\end{eqnarray}
These equations include mass transfers between fluids with $\mathbf{u}_{ij}^T$ being the velocity of the fluid transferred from $i$ to $j$ and $b_{ij}^T$ being the buoyancy of the fluid transferred from $i$ to $j$. A drag term, $\mathbf{D}_{ij}$ is also included to parameterise some of the missing terms because it is commonly used when conditional averaging is applied in fields such as multi-phase modelling \cite<eg>{GBB+07}. Sensitivity to $\mathbf{D}_{ij}$ will be investigated in section \ref{sec:results}.

@@

The instabilities reported by \citeA{WM19} showed growing divergence
in a fluid with vanishing volume fraction. \change[HW]{It is mass convergence
rather than velocity convergence that leads to pressure anomalies
that remove the convergence. Therefore if the volume (or mass) fraction
of fluid is vanishingly low then convergence cannot lead to an anomaly
in the total pressure.}{For single-fluid equations, divergence would lead to pressure anomalies which would act as a negative feedback on the divergence. However for multi-fluids if one of the volume fractions is too small then it cannot force pressure anomalies big enough to stabilise the divergence. Use of separate pressure for each fluid is proposed to ensure stability.}

We will assume that pressure in each fluid is given by a \remove[HW]{parameterised}
perturbation, $p_{i}$, from the total pressure, $P$:
\begin{equation}
P_{i}=P+p_{i}.
\end{equation}
\add[HW]{In order to derive a suitable form of parameterisation for $p_i$ we will first consider the} \citeA{BN86} \add{model of multi-phase flow.}

@@

\citeA{BN86} proposed a parameterisation for pressure differences
in a multi-phase explosion model that has been widely used in mathematics
and engineering for simulating, \add[HW]{for example}, fluidised beds of granular material \cite<e.g.>{EHM92} and two compressible fluids \cite<e.g.>{SA99}. \add[HW]{This model is not used here but is examined in order to derive a parameterisation of pressure differences between fluids suitable for atmospheric convection.}
\citeA{SA99} wrote transport equations for both the volume fraction,
$\sigma_{i}$, and for the mass fraction,
$\sigma_{i}\rho_{i}$  without transfers between fluids which in our notation are:
\begin{eqnarray}
\frac{\partial\sigma_{i}}{\partial t}+\mathbf{V}\cdot\nabla\sigma_{i} & = & \mu\left(P_{i}-P_{j}\right)
\label{eq:BNcompaction}
\\
\frac{\partial\sigma_{i}\rho_{i}}{\partial t}+\nabla\cdot\left(\sigma_{i}\rho_{i}\mathbf{u}_{i}\right) & = & 0
\label{eq:rhoTransport}
\end{eqnarray}
where $\mathbf{V}$ is the velocity of the interface between fluids,
$\mu$ is the compaction viscosity which controls how quickly the
pressure of each fluid relaxes to the mean pressure and index $j$
refers to the other fluid in a two fluid system. In our Boussinesq
system, $\rho_{i}$ does not appear in the continuity equation. Different
authors make different assumptions about the interface velocity but
if we assume that $\mathbf{V}=\sum_{i}\sigma_{i}\mathbf{u}_{i}$ and
$\rho_{i}$ is constant then eqns (\ref{eq:divFree}), (\ref{eq:sumOne}), (\ref{eq:BNcompaction}) and (\ref{eq:rhoTransport})
imply that $\mu p_{i}=-\sigma_{i}^{2}\nabla\cdot\mathbf{u}_{i}$.
If we assume that $\mathbf{V}=\mathbf{u}_{i}$ for each fluid then
we get $\mu p_{i}=-\sigma_{i}\nabla\cdot\mathbf{u}_{i}$.

@@

\citeA{TEB19} stabilised the multi-fluid equations by adding a diffusion
term $\nu\partial w^{2}/\partial z^{2}$ to the RHS of the $w$ equation.
If a fluid perturbation pressure is set to $-\nu\partial w/\partial z$
this is equivalent to the diffusion term in one dimension. However
adding a diffusion term will harm three dimensional simulations of
resolved convection in the free troposphere whereas adding a fluid
perturbation pressure will not. Inspired by both \citeA{BN86} and
\citeA{TEB19} we assume that the pressure in each fluid is given
by:
\begin{equation}
p_{i}=-\gamma\nabla\cdot(\sigma_i \mathbf{u}_{i})
\label{eq:Pi_div}
\end{equation}
where $\gamma$ is a \change[HW]{free coefficient}{compressibility with units $\text{m}^s \text{s}^{-1}$. 
This parameterisation of $p_i$ has $\sigma_i$ inside the divergence in order to ensure that $\sum_i p_i =0$ and therefore that $\sum_i P_i = P$ which makes the equations more striaghtforward to solve.}
\add[HW]{However it should be noted that this parameterisation destroys energy conservation.}

@@

Two test cases are developed to tune parameters and to \change[HW]{validate}{evaluate} the
use of \change[HW]{multi}{two}-fluid equations to represent dry, two-dimensional, sub-grid-scale
convection. The first is transient; the standard rising bubble of
\citeA{BF02}. \remove{We aim to reproduce the vertical heat transport, the
mean velocity of the rising air, the bouyancy of the rising air and
pressure differences between two fluids in a single column two fluid
model.} The second \remove[HW]{test case} is steady state with heat transfer at
the ground and uniform radiative cooling \add[HW]{(radiative convective equilibrium)}. For both test cases, the
reference solution is a single-fluid solution of the two-dimensional
Boussinesq equations at high horizontal and vertical resolution. \add[HW]{These will be compared with the two-fluid model run using one column. We aim to reproduce the vertical heat transport, the mean velocity of the rising air, the bouyancy of the rising air and the pressure differences between fluids in one column of the two-fluid model.}

>>>>

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%
\addeditor{HW}
\addeditor{WM}

@@

\citeA{TWV+18} proposed multi-fluid modelling of convection in order
to treat non-equilibrium and net mass transport \remove[HW]{correctly and} consistently.
\add[HW]{The multi-fluids approach has an advantage over ADHOC and CLUBB in solving prognostic equations for first-order moments directly (mean properties of eg. updraft and environment) rather than solving prognostic equations for higher-order moments of the whole fluid in order to diagnose means of updrafts and the environment. Solving directly for sub-fluid means should be less sensitive to closure assumptions.}
\remove{with the same numerics and physics to prognose properties of separate parts of the fluid, for example updrafts, the stable environment and downdrafts dynamics}. \citeA{Yano14} derived similar
equations \remove[HW]{for multiple segments} and showed how this is a generalisation
of the mass-flux formulation. The \citeA{TKP+18} ``extended eddy
diffusivity mass flux'' or extended EDMF scheme \remove[HW]{also} uses similar
equations but \citeA{TKP+18}
\change[HW]{didn't propose a suitable numerical method for solving the equations.}
{only solve the equations in a single column without coupling to the continuity equation of a dynamical core.}
\change[HW]{Multi-fluids promises big improvements
in the representation of sub-grid-scale convection and unification
with the representation of turbulence, particularly at the grey zone.
However, there is a long way to go, particularly as multi-fluids}
{The disadvantage of the multi-fluids approach is that it} requires
a dedicated model rather than being a stand alone parameterisation
for a dynamical core. \citeA{TEB19} and \citeA{WM19} presented solutions
of \change[HW]{the multi}{two}-fluid equations
\change[HW]{assuming two fluids that share}{using} the same pressure
\add[HW]{for both fluids}.
\remove[HW]{(or that pressure differences are predominantly due to drag
between the fluids)}. \citeA{TEB19} showed that these equations \remove[HW]{(without
drag)} are in fact unstable and can be stabilised by diffusion of vertical
velocity whereas \citeA{WM19} stabilised the equations with diffusion
and drag between fluids. Neither of these stabilisation techniques
are acceptable;
\add[HW]{sufficient diffusion to stabilise the multi-fluid equations may smooth out real features of the flow and the diffusion between fluids and drag used by \protect\citeA{WM19} meant that both fluids tended to move together rather than through each other.}
In this paper we show that there are significant
and sustained pressure differences between the fluids (section \ref{sec:results}) \add[HW]{and that the use of separate pressures stabilises the equations}.

@@

Many of the challenges of representing convection with the multi-fluid
Navier-Stokes equations \cite<e.g. those described in>{WM19} carry
over to the Boussinesq equations \add[HW]{because} convection is buoyancy dominated
and close to incompressible. The resolved test cases presented in
section \ref{sec:results} give very similar results for Navier-Stokes
and Buossinesq equations. Therefore\change[WM]{ for simplicity we focus on the}{, the}
\remove[HW]{multi-fluid} Boussinesq equations without background stratification \add[WM]{are chosen for simplicity}:
\note[HW]{This is additional material until the start of 2.1}
\begin{eqnarray}
\frac{\partial \mathbf{u}}{\partial t} + \nabla\cdot(\mathbf{u}\mathbf{u})
+\nabla P & = & b \mathbf{k}+\nu\nabla^{2}\mathbf{u}
\label{eq:singleMom}
\\
\frac{\partial b}{\partial t} + \nabla\cdot(\mathbf{u}b)
 & = & \alpha\nabla^{2}b
\label{eq:singleb}
\\
\nabla\cdot\mathbf{u} & = & 0
\label{eq:singleDivFree}
\end{eqnarray}
where variables are defined in table \ref{tab:defns}.

\begin{table}
\begin{tabular}{c>{\raggedright}p{0.85\textwidth}}
$\sigma_{i}$ & Volume fraction of fluid $i$ so that $\sum_{i}\sigma_{i}=1$
\tabularnewline
\add[HW]{$\mathbf{u}$} & 
\add[HW]{Velocity of the single-fluid equations  ($\text{m}\text{s}^{-1}$) and velocity averaged over all fluids of the multi-fluid equations}
\tabularnewline
$\mathbf{u}_{i}$ & Velocity of fluid $i$ ($\text{m}\text{s}^{-1}$)
\tabularnewline
$w_{i}$ & Vertical component of $\mathbf{u}_{i}$ ($\text{m}\text{s}^{-1}$)\tabularnewline
$S_{ij}$ & Transfer from fluid $i$ to $j$ ($\text{s}^{-1}$)
\tabularnewline
\add[HW]{$P$} & \add[HW]{Non-hydrostatic pressure of the single-fluid equations $=p^{\prime}/\rho_{r}$ ($\text{m}^{2}\text{s}^{-2}$) and pressure averaged over all fluids of the multi-fluid equations}
\tabularnewline
$P_i$ & Non-hydrostatic pressure of fluid $i$
\tabularnewline
\add[HW]{$p_i$} & \add[HW]{$P_i - P$}
\tabularnewline
$\psi_{ij}^{T}$ & Value of variable $\psi$ transferred from fluid $i$ to $j$
\tabularnewline
$b$ & Buoyancy of single-fluid equations defined as -$g\rho^{\prime}/\text{\ensuremath{\rho}}_{r}$
where $\rho^{\prime}$ are departures in density from a horizontally
uniform reference, $\rho_{r}$ ($\text{m}\text{s}^{-2}$) \add[HW]{and buoyancy averaged over all fluids of the multi-fluid equations}
\tabularnewline
$b_{i}$ & Buoyancy of fluid $i$ ($\text{m}\text{s}^{-2}$)
\tabularnewline
$D_{i}\big/Dt$ & Total derivative with respect to fluid $i$ $=\partial/\partial t+\mathbf{u}_{i}\cdot\nabla$
\tabularnewline
$\mathbf{D}_{ij}$ & Drag on fluid $i$ from fluid $j$ ($\text{m}\text{s}^{-2}$)
\tabularnewline
$\gamma$ & \change[HW]{Coefficient for setting the pressure local for each fluid}{Compressibility for calculating $p_i$} ($\text{m}^{2}\text{s}^{-1}$)
\tabularnewline
$C_{D}$ & Drag coefficient $=0$ or 1
\tabularnewline
$r_{c}$ & Plume radius used for defining the drag between fluids (m)\tabularnewline
$\alpha$ & Diffusivity of buoyancy ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
$\nu$ & Viscosity ($\text{m}^{2}\text{s}^{-1}$)
\tabularnewline
\add[WM]{$\mathbf{k}$} & \add[WM]{Unit vector in the $z$-direction.}
\tabularnewline
\end{tabular}
\caption{Definitions of variables for the single- and multi-fluid Boussinesq equations.\label{tab:defns}}
\end{table}

The single-fluid Boussinesq equations will be conditionally averaged following \citeA{TWV+18} to give the multi-fluid Boussinesq equations. Lagrangian labels for each fluid $i$ are defined as $I_i=1$ in fluid $i$ and zero elsewhere. 
For the derivation of the conditionally averaged equations \change[WM]{we will not consider relabelling of fluid}{the relabelling of fluids will not be considered,} \add[HW]{and so} $I_i$ satisfies:
\begin{equation}
\frac{DI_i}{Dt}=0.
\label{eq:LagrangianLabel}
\end{equation}
\change[WM]{Applying conditional volume average to $I_i$ we define the volume fraction of fluid $i$ and the velocity and buoyancy of fluid $i$}{By applying a conditional volume average (denoted by the overbar), the fluid $i$ volume fractions, velocities and buoyancies are defined by}:
\begin{eqnarray}
\sigma_i &=& \overline{I_i} \\
\sigma_i \mathbf{u}_i &=& \overline{I_i \mathbf{u}} \\
\sigma_i b_i &=& \overline{I_i b}
\label{eq:defineFluidFields}
\end{eqnarray}
which implies
\begin{equation}
\sum_{i}\sigma_{i}  =  1.
\label{eq:sumOne}
\end{equation}
\change[WM]{As we are considering the Boussinesq equations we assume}{As the system is governed by the Boussinesq equations, it is assumed} that variations in density only affect the \change[WM]{buoyancy terms}{definition of the buoyancy terms (table \protect\ref{tab:defns})} so $\mathbf{u}_i$ and $b_i$ are defined without density weighting.

To derive a transport equation for $\sigma_i$\change[WM]{ we combine (\protect\ref{eq:LagrangianLabel})}{, (\protect\ref{eq:LagrangianLabel}) is combined} with $I_i$ times (\ref{eq:singleDivFree}) and volume average. After permuting operators with the averaging \cite<as described by>{TWV+18}\add[WM]{,} and using the definitions of $\sigma_i$ and $\mathbf{u}_i$\add[WM]{,} \remove[WM]{we get} the evolution equation for $\sigma_i$\add[WM]{ is obtained}:
\begin{equation}
\frac{\partial\sigma_{i}}{\partial t}+\nabla\cdot\sigma_{i}\mathbf{u}_{i}  =  \sum_{j\ne i}\left\{ \sigma_{j}S_{ji}-\sigma_{i}S_{ij}\right\}
\label{eq:sigma}
\end{equation}
which also includes transfers between fluids with $\sigma_{i}S_{ij}$ being the transfer from fluid $i$ to $j$.

Next \change[WM]{we will consider conditionally averaging the terms of }{the conditional averaging of terms in} the momentum (\ref{eq:singleMom}) and buoyancy (\ref{eq:singleb}) equations\add[WM]{ is considered}. Conditionally averaging the advection terms leads to sub-filter-scale fluxes, $F_{SF}$, since the velocity and buoyancy are not uniform within each fluid at the filter scale:
\begin{eqnarray}
\overline{I_i \nabla \cdot (\mathbf{u} \mathbf{u})} &=& \nabla \cdot (\sigma_i \mathbf{u}_i  \mathbf{u}_i) + F_\text{SF}^{\mathbf{u}_i} \\
\overline{I_i \nabla (\mathbf{u} b)} &=&  \nabla \cdot (\sigma_i \mathbf{u}_i b_i) + F_\text{SF}^{b_i}
\label{eq:filterAdvection}
\end{eqnarray}
The sub-filter-scale fluxes will be absorbed into the diffusion terms for momentum and buoyancy.
%
\note[HW]{Dan, can we do better than this?}
\change[WM]{Lots of}{Many} terms appear when attempting to conditionally average the diffusion terms which we do not know how to parameterise. \change[WM]{Therefore we will use simply}{The conditionally averaged diffusion terms are therefore assumed to behave like}:
\begin{eqnarray}
\overline{I_i \nu \nabla^2 \mathbf{u}} &=& \sigma_i \nu \nabla^2 \mathbf{u}_i \\
\overline{I_i \alpha \nabla^2 b} &=& \sigma_i \alpha \nabla^2b_i .
\end{eqnarray} 
%
\remove[WM]{Unlike \protect\citeA{TWV+18}} \change[WM]{we define separate pressures for each fluid, $P_i$. The pressures are defined to satisfy:}{The pressures for each fluid, $P_i$, are defined to satisfy:}
\begin{equation}
\overline{I_i \nabla P} = \sigma_i \nabla P_i
\end{equation}
\add[WM]{which differs from \protect\citeA{TWV+18} who use a shared fluid pressure.}
%
Combining \change[WM]{these}{the above terms} gives the conditionally averaged momentum and buoyancy equations:
\begin{eqnarray}
\frac{\partial\sigma_i\mathbf{u}_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} \mathbf{u}_{i})
+ \sigma_i \nabla P_{i}
& = &
\sigma_i b_{i}\mathbf{k}
+
\nu\sigma_i\nabla^{2}\mathbf{u}_{i}
+
\sum_{j\ne i}\left\{ \sigma_{j}S_{ji}\mathbf{u}_{ji}^{T}
 - \sigma_{i}\mathbf{D}_{ij}\right\}
\label{eq:momFlux}\\
\frac{\partial\sigma_i b_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} b_{i})
& = &
\alpha\sigma_i\nabla^{2}b_{i}
+
\sum_{j\ne i} \sigma_{j}S_{ji} b_{ji}^{T}.
\label{eq:bFlux}
\end{eqnarray}
These equations include mass transfers between fluids with $\mathbf{u}_{ij}^T$ being the velocity of the fluid transferred from $i$ to $j$ and $b_{ij}^T$ being the buoyancy of the fluid transferred from $i$ to $j$. A drag term, $\mathbf{D}_{ij}$ is also included to parameterise some of the missing terms because it is commonly used when conditional averaging is applied in fields such as multi-phase modelling \cite<eg>{GBB+07}. Sensitivity to $\mathbf{D}_{ij}$ will be investigated in section \ref{sec:results}.

@@

The instabilities reported by \citeA{WM19} showed growing divergence
in a fluid with vanishing volume fraction. \change[HW]{It is mass convergence
rather than velocity convergence that leads to pressure anomalies
that remove the convergence. Therefore if the volume (or mass) fraction
of fluid is vanishingly low then convergence cannot lead to an anomaly
in the total pressure.}{For single-fluid equations, divergence would lead to pressure anomalies which would act as a negative feedback on the divergence. However for multi-fluids if one of the volume fractions is too small then it cannot force pressure anomalies big enough to stabilise the divergence. Use of a separate pressure for each fluid is therefore proposed to ensure stability.}

\change[WM]{We will assume that}{It is assumed that the} pressure in each fluid is given by a \remove[HW]{parameterised}
perturbation, $p_{i}$, from the total pressure, $P$:
\begin{equation}
P_{i}=P+p_{i}.
\end{equation}
\add[HW]{In order to derive a suitable form of parameterisation for $p_i$ we will first consider the} \citeA{BN86} \add{model of multi-phase flow.}

@@

\citeA{BN86} proposed a parameterisation for pressure differences
in a multi-phase explosion model that has been widely used in mathematics
and engineering for simulating, \add[HW]{for example}, fluidised beds of granular material \cite<e.g.>{EHM92} and two compressible fluids \cite<e.g.>{SA99}. \add[HW]{This model is not used here but is examined in order to derive a parameterisation of pressure differences between fluids suitable for atmospheric convection.}
\citeA{SA99} wrote transport equations for both the volume fraction,
$\sigma_{i}$, and for the mass fraction,
$\sigma_{i}\rho_{i}$  without transfers between fluids which in our notation are:
\begin{eqnarray}
\frac{\partial\sigma_{i}}{\partial t}+\mathbf{V}\cdot\nabla\sigma_{i} & = & \mu\left(P_{i}-P_{j}\right)
\label{eq:BNcompaction}
\\
\frac{\partial\sigma_{i}\rho_{i}}{\partial t}+\nabla\cdot\left(\sigma_{i}\rho_{i}\mathbf{u}_{i}\right) & = & 0
\label{eq:rhoTransport}
\end{eqnarray}
where $\mathbf{V}$ is the velocity of the interface between fluids,
$\mu$ is the compaction viscosity which controls how quickly the
pressure of each fluid relaxes to the mean pressure and index $j$
refers to the other fluid in a two fluid system. In our Boussinesq
system, $\rho_{i}$ does not appear in the continuity equation. Different
authors make different assumptions about the interface velocity but
if \change[WM]{we assume}{one assumes} that $\mathbf{V}=\sum_{i}\sigma_{i}\mathbf{u}_{i}$ and
$\rho_{i}$ is constant then eqns (\ref{eq:divFree}), (\ref{eq:sumOne}), (\ref{eq:BNcompaction}) and (\ref{eq:rhoTransport})
imply that $\mu p_{i}=-\sigma_{i}^{2}\nabla\cdot\mathbf{u}_{i}$.
\change[WM]{If we assume that $\mathbf{V}=\mathbf{u}_{i}$ for each fluid then
we get $\mu p_{i}=-\sigma_{i}\nabla\cdot\mathbf{u}_{i}$.}{Instead assuming $\mathbf{V}=\mathbf{u}_{i}$ for each fluid results in $\mu p_{i}=-\sigma_{i}\nabla\cdot\mathbf{u}_{i}$.}

@@

\citeA{TEB19} stabilised the multi-fluid equations by adding a diffusion
term $\nu\partial w^{2}/\partial z^{2}$ to the RHS of the $w$ equation.
If a fluid perturbation pressure is set to \change[WM]{$-\nu\partial w/\partial z$}{$p_i=-\nu\partial w/\partial z$}
\add[WM]{then} this is equivalent to the diffusion term in one dimension. However
adding a diffusion term will harm three dimensional simulations of
resolved convection in the free troposphere whereas adding a fluid
perturbation pressure will not. Inspired by both \citeA{BN86} and
\citeA{TEB19}\change[WM]{ we assume that the pressure in each fluid is given
by}{, the pessure in each fluid is chosen to be}:
\begin{equation}
p_{i}=-\gamma\nabla\cdot(\sigma_i \mathbf{u}_{i})
\label{eq:Pi_div}
\end{equation}
where $\gamma$ is a \change[HW]{free coefficient}{compressibility with units $\text{m}^s \text{s}^{-1}$. 
This parameterisation of $p_i$ has $\sigma_i$ inside the divergence in order to ensure that $\sum_i p_i =0$ and therefore that $\sum_i P_i = P$ which makes the equations more striaghtforward to solve.}
\add[HW]{However it should be noted that this parameterisation destroys energy conservation.}

@@

Two test cases are developed to tune parameters and to \change[HW]{validate}{evaluate} the
use of \change[HW]{multi}{two}-fluid equations to represent dry, two-dimensional, sub-grid-scale
convection. The first is transient; the standard rising bubble of
\citeA{BF02}. \remove{We aim to reproduce the vertical heat transport, the
mean velocity of the rising air, the bouyancy of the rising air and
pressure differences between two fluids in a single column two fluid
model.} The second \remove[HW]{test case} is steady state with heat transfer at
the ground and uniform radiative cooling \add[HW]{(radiative convective equilibrium)}. For both test cases, the
reference solution is a single-fluid solution of the two-dimensional
Boussinesq equations at high horizontal and vertical resolution. \add[HW]{These will be compared with the two-fluid model run using one column.} \add[WM]{The aim is to reproduce} \add[HW]{the vertical heat transport, the mean velocity of the rising air, the bouyancy of the rising air and the pressure differences between fluids in one column of the two-fluid model.}
